---
title: "Example R Markdown with Plots"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Required Libraries

```{r load-libraries}
library(ggplot2)
library(dplyr)
library(ape)
library(caper)
library(nlme)
library(geiger)
library(phytools)
```

## Example Data
We will use the built-in `mtcars` dataset.

```{r display-data}
head(mtcars)
```

## Base R Plots

### Scatter Plot
This plot is useful for visualizing relationships between two continuous variables.
```{r base-scatter}
plot(mtcars$mpg, mtcars$hp, 
     main = "MPG vs HP", 
     xlab = "Miles Per Gallon", 
     ylab = "Horsepower", 
     pch = 19, col = "blue")
```

### Histogram
A histogram is ideal for understanding the distribution of a single continuous variable.
```{r base-hist}
hist(mtcars$mpg, 
     main = "Histogram of MPG", 
     xlab = "Miles Per Gallon", 
     col = "lightblue", 
     border = "black")
```

## ggplot2 Visualizations

### ggplot Scatter Plot
This scatter plot is used to analyze the correlation between two continuous variables, such as fuel efficiency and horsepower.
```{r ggplot-scatter}
ggplot(mtcars, aes(x = mpg, y = hp)) +
  geom_point(color = "blue", size = 3) +
  labs(title = "MPG vs HP", x = "Miles Per Gallon", y = "Horsepower") +
  theme_minimal()
```

### ggplot Histogram
Histograms help in identifying the distribution and frequency of data points within specified bins.
```{r ggplot-hist}
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram(binwidth = 2, fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Histogram of MPG", x = "Miles Per Gallon", y = "Count") +
  theme_minimal()
```

### ggplot Boxplot
Boxplots are useful for comparing distributions across categorical variables and detecting outliers.
```{r ggplot-boxplot}
ggplot(mtcars, aes(x = factor(cyl), y = mpg, fill = factor(cyl))) +
  geom_boxplot() +
  labs(title = "MPG by Number of Cylinders", x = "Cylinders", y = "Miles Per Gallon") +
  theme_minimal()
```

### ggplot Bar Plot
Bar plots are used to visualize categorical data by showing the frequency of each category.
```{r ggplot-barplot}
ggplot(mtcars, aes(x = factor(cyl))) +
  geom_bar(fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Count of Cars by Cylinder", x = "Cylinders", y = "Count") +
  theme_minimal()
```

## Phylogenetic Tree Plot
Phylogenetic trees are used to visualize evolutionary relationships between species or other hierarchical structures.
```{r phylo-tree}
tree <- read.tree("sampleTree.nwk")
plot(tree, main = "Phylogenetic Tree", cex = 0.1)
```

## Phylogenetic Generalized Least Squares (PGLS)
PGLS is used to account for phylogenetic relatedness in regression models. 
Here, we will perform a PGLS analysis using the `caper` package.

```{r pgls-analysis}
data <- read.csv("sampleData.csv")

#correlation strucutre
bm<-corBrownian(1,tree)
model1<-gls(NeoplasiaPrevalence~max_longevity, data=data, correlation=bm)

```

Plot challenge

### Dataset 1: Time Series Data
This dataset represents a time series with a continuous trend.

```{r time-series-data}
time_series_data <- data.frame(
  time = 1:100,
  value = cumsum(rnorm(100))
)
head(time_series_data)
```

### Dataset 2: Categorical Comparison
This dataset represents counts for different categories.

```{r categorical-data}
categorical_data <- data.frame(
  category = rep(c("A", "B", "C", "D"), each = 10),
  value = rpois(40, lambda = 10)
)
head(categorical_data)
```

### Dataset 3: Correlation Data
This dataset represents a relationship between two continuous variables.

```{r correlation-data}
correlation_data <- data.frame(
  x = rnorm(100, mean = 50, sd = 10),
  y = rnorm(100, mean = 100, sd = 20)
)
head(correlation_data)
```



